Does at-issueness predict projection? Further investigations of the Gradient Projection Principle

### 2022-10-27 writing/investigating-GPP.tex
worked on section 2 and appendices

@@

Exps.~1-3 were designed to investigate whether not-at-issueness and projection are positively correlated, as predicted by the Gradient Projection Principle, to compare the projection of contents from different entailment-canceling environments, and to compare different measures of at-issueness.\footnote{\label{f-github}The experiments, data, and R code for generating the figures and analyses of the experiments reported in this paper are available at \url{https://github.com/judith-tonhauser/projective-probability}. All experiments were conducted with IRB approval. Exps.~1b, 2b, and 3b were preregistered: \url{https://osf.io/cxq47}.} To be able to compare across the experiments, the experiments investigated the same contents and each of the experiments consisted of two blocks: a projection block, in which the projection of the contents was measured, and an at-issueness block, in which the at-issueness of the contents was measured. The at-issueness blocks of Exps.~1-3 implemented different measures of at-issueness, which allows for a comparison of at-issueness measures. To compare projection from different entailment-canceling environments, each of Exps.~1-3 consisted of four experiments (for a total of 12 experiments) where the contents were realized by expressions in a polar question, in the scope of sentential negation, in the scope of the epistemic possibility modal {\em perhaps}, or in the antecedent of a conditional. For an overview of the 12 experiments see Table \ref{t:overview}: the name of each of the 12 experiments identifies whether belongs to the set of Exps.~1, 2, or 3 (which differ in at-issueness measure) and whether the embedding-environment is a polar question (`q'), negation (`n'), the epistemic possibility modal (`m'), or the antecedent of a conditional (`c').

Projection was measured across all 12 experiments with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017,lorson2018,tbd-variability,mahler-nels,mahler2020,demarneffe-etal-sub23}).\footnote{For other diagnostics for projection see, e.g.\ \citealt{smith-hall11,xue-onea11}, and \citealt{brst-lang11}; see also the discussion in \citealt{tbd-variability}.}  On this diagnostic, participants are presented with an utterance, like one of Daniel's utterances in (\ref{proj-stim}), in which the content to be diagnosed for projection (here, that Julian dances salsa) occurs in one of the four entailment-canceling environments:

\begin{exe}
\ex\label{proj-stim} `certain that' diagnostic for projection
\begin{xlist}
\ex Daniel: Did Cole discover that Julian dances salsa? \hfill [Exps.~1q, 2q, 3q]
\ex Daniel: Cole didn't discover that Julian dances salsa. \hfill [Exps.~1n, 2n, 3n]
\ex Daniel: Perhaps Cole discovered that Julian dances salsa. \hfill [Exps.~1m, 2m, 3m]
\ex Daniel: If Cole discovered that Julian dances salsa, Logan will be joyful. \hfill [Exps.~1c, 2c, 3c]
\end{xlist}
Question to participants: Is Daniel certain that Julian dances salsa?
\end{exe}
To measure projection, participants are asked whether the speaker is certain of the relevant content. In \ref{proj-stim}, for instance, participants are asked whether Daniel is certain that Julian dances salsa. If a participant takes the speaker to be certain of the content, we assume that the content projects out of the entailment-canceling environment; if a participant does not take the speaker to be certain of the content, we assume that the content does not project.

At-issueness was measured with different diagnostics across Exps.~1-3. In Exps.~1, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c.

@@

We recruited 250-300 participants for each of Exps.~1-3. Participants for Exp.~1q were recruited on Amazon's Mechanical Turk platform; the participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs. Participants for the remaining experiments were recruited on Prolific; these participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%. Information on recruited participants (total number, ages, gender) and the payments they received can be found in Supplement \ref{a-participants}.

@@

The projection and at-issueness of the contents of 20 clauses (such as {\em Julian dances salsa}) were measured in separate blocks in Exps.~1-3. For the full set of clauses, see Appendix \ref{a-clauses}.

In the projection blocks of Exps.~1-3, target stimuli consisted of a polar question sentence, a negated sentence, a sentence with {\em perhaps}, or a conditional sentence that was uttered by a named speaker, as shown in Fig.~\ref{fig:projection-trials}. Each sentence was formed by realizing the 20 clauses as the complements of the 20 clause-embedding predicates in Fig.~\ref{fig:predicates}.\footnote{The proper names that realized the speakers, the subjects of the clause-embedding predicates, the subjects of the complement clauses, and the subjects of the consequents  were all unique. The indirect object of {\em inform} was realized by the proper name {\em Sam}. The predicates differed in the tense in which they were realized. In polar interrogatives as well as in negated and modalized declaratives, eventive predicates, like {\em discover} and {\em hear}, were realized in the past tense and stative predicates, like {\em know} and {\em be annoyed}, were realized in the present tense.  In the antecedents of conditionals, all predicates were realized in the present tense. For the consequents of the conditional sentences see Supplement \ref{a-target}.} Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody ask somebody else a question. Projection was measured using the `certain that' diagnostic \cite{djaerv-bacovcin-salt27,tbd-variability, lorson2018, mahler2020}: participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact. They gave their responses on a slider marked `no' at one end (coded as 0) and `yes' at the other (coded as 1). Greater speaker commitment to the CC should result in higher slider ratings.

@@

In the at-issueness blocks of Exps.~1, target

In the projection blocks of Exps.~1-3, target stimuli consisted of a polar question sentence, a negated sentence, a sentence with {\em perhaps}, or a conditional sentence that was uttered by a named speaker, as shown in Fig.~\ref{fig:projection-trials}. Each sentence was formed by realizing the 20 clauses as the complements of the 20 clause-embedding predicates in Fig.~\ref{fig:predicates}.\footnote{The proper names that realized the speakers, the subjects of the clause-embedding predicates, the subjects of the complement clauses, and the subjects of the consequents  were all unique. The indirect object of {\em inform} was realized by the proper name {\em Sam}. The predicates differed in the tense in which they were realized. In polar interrogatives as well as in negated and modalized declaratives, eventive predicates, like {\em discover} and {\em hear}, were realized in the past tense and stative predicates, like {\em know} and {\em be annoyed}, were realized in the present tense.  In the antecedents of conditionals, all predicates were realized in the present tense. For the consequents of the conditional sentences see Supplement \ref{a-target}.} Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody ask somebody else a question. Projection was measured using the `certain that' diagnostic \cite{djaerv-bacovcin-salt27,tbd-variability, lorson2018, mahler2020}: participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact. They gave their responses on a slider marked `no' at one end (coded as 0) and `yes' at the other (coded as 1). Greater speaker commitment to the CC should result in higher slider ratings.

The at-issueness diagnostics differed between Exps.~1, 2, and 3: at-issueness was measured in Exp.~1q, 2q, and 3q, with variants of the question-based diagnostic of at-issueness, namely the `asking whether' diagnostic in Exp.~1q, and the `yes, $p$' diagnostic in Exps.~2q and 3q, as shown in Figure \ref{fig-ai-trials-q}.

, with the `sure that' diagnostic in Exps.~1n, 1m, and 1c, with the `yes' diagnostic in Exps.~2q and 3q, with the `assent with positive continuation' diagnostic in Exps.~2n, 2m, and 2c,  and with the `assent with adversative continuation' diagnostic in Exps.~3n, 3m, and 3c.

To assess whether participants were attending to the task, each experiment also included six control stimuli, which were also utterances made by a named speaker. (The controls are provided in Supplement \ref{a-control}.) Each participant saw a random set of 26 stimuli: each set contained one target utterance for each of the 20 clause-embedding predicates (each with a unique complement clause) and the same 6 control stimuli. Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block. Block order and within-block trial order was randomized.

@@

\caption{Target trials in at-issueness blocks of Exps.~2 for the complement {\em Julian dances salsa}.}\label{fig:ai-exps3}
\end{figure}

@@

====

The projection block also included 6 control trials, which functioned as attention checks. The content of these stimuli was expected not to project: For example, in Fig.~\ref{fig-exp1-projection-control}, the speaker is not committed to the main clause content, that Samantha has a new hat. The same 6 main clauses were also used to form 6 filler trials in the prior block; a sample stimulus is given in Fig.~\ref{fig-exp1-prior-filler}. These filler stimuli were not used to assess participants' attention. For the full set of stimuli see Supplement \ref{a-stim}.

Each participant's stimulus set was semi-randomly generated by first randomly pairing up the 20 predicates and clauses. Half of the stimuli were then randomly assigned the respective clause's higher-probability fact, and half its lower-probability fact. Participants completed a total of 52 trials: 20 target trials in each block, 6 control trials in the projection block, and 6 filler trials in the prior block. Each participant completed the same 6 filler and control trials. Block order and within-block trial order were randomized.

After completing the experiment, participants filled out a short optional demographic survey. To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.

@@

\section{20 complement clauses}\label{a-clauses}

The following clauses realized the complements of the predicates in Exps.~1-3:

@@

\section{Consequents for conditional target stimuli in Exps.~1c, 2c, and 3c}\label{a-target}

@@

The control stimuli in Exps.~1-3 were main clauses. In the question embedding experiments (Exps.~1q, 2q and 3q), the control stimuli consisted of the polar questions in (1). The NRRCs in parentheses were realized in Exps.~2q and 3q, where at-issueness was measured with an assent diagnostic: the control stimuli here consisted of two clauses (like the target stimuli), allowing the relevant speaker to assent with one of the two clauses.

@@

We expected participants to give low responses on the `certain that' diagnostic for the control stimuli in (1), indicating that the speaker is not certain of the main clause content, because main clause content is hypothesized to not project out of the question. These expectations were borne out, as shown in the third column of Table~\ref{t-controls} for Exps.~1q, 2q, and 3q. We also expected participants to give low responses on the at-issueness diagnostics for the control stimuli in (1), indicating that the main clause content is at-issue. These expectations were borne out for Exps.~1q and 2q, as shown in the fourth column of Table~\ref{t-controls}, but not for Exp.~3q.

In the remaining experiments, the control stimuli consisted of the positive declarative variants of (1) given in (2). The NRRCs in parentheses were realized in Exps.~2 and 3 for the reason explained above, as well as in Exp.~1c, to make the control stimuli more similar to the target stimuli (which also consisted of two clauses: the antecedent and the consequent).

@@

We expected participants to give high responses on the `certain that' diagnostic for the control stimuli in (2), indicating that the speaker is certain of the main clause content, because speakers are hypothesized to be committed to the asserted main clause content. These expectations were borne out, as shown in the third column of Table~\ref{t-controls} for the `n', `m' and `c' variants of Exps.~1, 2, and 3. We expected participants to give low responses on the at-issueness diagnostics for the control stimuli in (2), indicating that the main clause content is at-issue. These expectations were borne out for the `n', `m', and `c' variants of Exps.~1, as shown in the fourth column of Table~\ref{t-controls}, but not for the `n', `m', and `c' variants of Exps.~2 and 3.

\begin{table}[h!]
\centering
\begin{tabular}{r r r r l }
& &  \multicolumn{2}{c}{Mean ratings} &  \\ 
Exp. & Control stimuli & Certainty & At-issueness & At-issueness measure \\ 
\hline
1q & (1) & .14 & .05  & asking whether $c$ \\
1n & (2) &  .95 & .04 & sure that $c$\\
1m & (2) & .96 & .03 & sure that $c$\\
1c & (2) with NRRC & .94  & .08 & sure that $c$\\
\hline
2q & (1) with NRRC & .18 & .07 & {\em yes}, $c$\\
2n & (2) with NRRC& .96 & .22 & {\em yes, that's true}, $c$\\
2m & (2) with NRRC& .96 & .25 & {\em yes, that's true}, $c$\\
2c & (2) with NRRC& .96 & .22 & {\em yes, that's true}, $c$\\
\hline
3q & (1) with NRRC & .17 &  .28 & {\em yes}, but $\neg c'$\\
3n & (2) with NRRC & .94 & .44 & {\em yes, that's true}, but $\neg c'$\\
3m & (2) with NRRC & .93 & .50 & {\em yes, that's true}, but $\neg c'$\\
3c & (2) with NRRC & .93 & .53 & {\em yes, that's true}, but $\neg c'$\\
\hline
\end{tabular}
\caption{Mean certainty and at-issueness ratings for control stimuli, for self-declared American English participants}\label{t-controls}
\end{table}

@@

\item `controls': Participants' data were excluded if their mean rating on the 6 main clause control items in the projection block was more than 2 sd above the group mean (in Exps.~1q, 2q, and 3q) or more than 2 sd below the group mean (in the remaining experiments). Participants' data were also excluded if their mean rating on the 6 main clause control items in the at-issueness block was more than 2 sd above the group mean (across all experiments).

@@

\section{Comparing projection with different entailment-canceling environments}\label{a-projection}

>>>>

\subsection{At-issueness diagnostics}\label{s:ai-diagnostics}

In Exps.~1, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q, illustrated in (\ref{ai-exps1}a), and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c, illustrated in (\ref{ai-exps1}b). The `asking whether' diagnostic requires that the content to be diagnosed is realized by an expression in a polar question; this diagnostic was also used in \citealt{tbd-variability} Exps.~1. A speaker (Christopher in (\ref{ai-exps1}a) asks a polar question and the participant is asked whether the speaker is asking about the content to be diagnosed (here, that Julian dances salsa). If a participant takes the speaker to be asking of the content, the content is assumed to be at-issue; if a participant does not take the speaker to be asking about the content, the content is assumed to be not-at-issue. The `sure that' diagnostic requires that the content to be diagnosed is realized by an expression in declarative sentence; this diagnostic was also used in \citealt{tbd-variability} Exps.~2. A speaker utters the declarative sentence and the addressee (here, Sandy) asks {\em Are you sure?}. The speaker response with {\em Yes, I am sure that} followed by the expression that denotes the content to be diagnosed. Participants are asked whether the speaker answered the addressee's question. If a participant takes the speaker to have answered the question, the content is assumed to be at-issue; if a participant does not take the speaker to have answered the question, the content is assumed to be not-at-issue. {\bf move all this into section 1}

@@

Exps.~1-3 were designed to investigate the three research questions in (\ref{rqs}), repeated here for convenience:\footnote{\label{f-github}The experiments, data, and R code for generating the figures and analyses of the experiments reported in this paper are available at \url{https://github.com/judith-tonhauser/projective-probability}. All experiments were conducted with IRB approval. Exps.~1b, 2b, and 3b were preregistered: \url{https://osf.io/cxq47}.}

\begin{exe}
\ex\label{rqs} Research questions
\begin{xlist}
\ex Does the projection of content differ across entailment-canceling environments?
\ex Are the results of different measures of at-issueness positively correlated?
\ex Is the projection of content positively correlated with its not-at-issueness, as predicted by \citepos{tbd-variability} Gradient Projection Principle?
\end{xlist}
\end{exe}
To address these questions, Exps.~1-3 investigated the projection and at-issueness of same 20 contents. Each of the experiments consisted of two blocks: a projection block, in which the projection of the contents was measured, and an at-issueness block, in which the at-issueness of the contents was measured. To address the research question in (\ref{rqs}b), different measures of at-issueness were implemented in the at-issueness blocks of Exps.~1-3, respectively. To address the research question in (\ref{rqs}a), each of Exps.~1-3 consisted of four experiments (for a total of 12 experiments) where the contents were realized by expressions in a polar question, in the scope of sentential negation, in the scope of the epistemic possibility modal {\em perhaps}, or in the antecedent of a conditional. The research question (\ref{rqs}c) is addressed by comparing the projection and at-issueness ratings of each of the 12 experiments. For an overview of the 12 experiments see Table \ref{t:overview}, where the name of each experiment identifies whether belongs to the set of Exps.~1, 2, or 3 (which differ in at-issueness measures) and whether the embedding-environment is a polar question (`q'), negation (`n'), the epistemic possibility modal (`m'), or the antecedent of a conditional (`c').

Projection was measured across all 12 experiments with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017,lorson2018,tbd-variability,mahler-nels,mahler2020,demarneffe-etal-sub23}).\footnote{For other diagnostics for projection see, e.g.\ \citealt{smith-hall11,xue-onea11}, and \citealt{brst-lang11}; see also the discussion in \citealt{tbd-variability}.}  On this diagnostic, participants are presented with an utterance, like one of Daniel's utterances in (\ref{proj-stim}), in which the content to be diagnosed for projection (here, that Julian dances salsa) is realized by an expression that occurs in one of the four entailment-canceling environments, that is, in a polar question (\ref{proj-stim}a), in the scope of negation (\ref{proj-stim}b), in the scope of the epistemic possibility modal {\em perhaps}, or in the antecedent of a conditional.

\begin{exe}
\ex\label{proj-stim} `certain that' diagnostic for projection
\begin{xlist}
\ex Daniel: Did Cole discover that Julian dances salsa? \hfill [Exps.~1q, 2q, 3q]
\ex Daniel: Cole didn't discover that Julian dances salsa. \hfill [Exps.~1n, 2n, 3n]
\ex Daniel: Perhaps Cole discovered that Julian dances salsa. \hfill [Exps.~1m, 2m, 3m]
\ex Daniel: If Cole discovered that Julian dances salsa, Logan will be joyful. \hfill [Exps.~1c, 2c, 3c]
\end{xlist}
Question to participants: Is Daniel certain that Julian dances salsa?
\end{exe}
To measure projection, participants are asked whether the speaker is certain of the relevant content. In \ref{proj-stim}, for instance, participants are asked whether Daniel is certain that Julian dances salsa. If a participant takes the speaker to be certain of the content, the content is assumed to project out of the entailment-canceling environment; if a participant does not take the speaker to be certain of the content, the content is assumed to not project.

As shown in Table \ref{t:overview}, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q (as this diagnostic can only applied to content in polar questions) and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c (as this diagnostic can only be applied to content in declarative sentences). As discussed in section \ref{s:ai-diagnostics}, these two measures of at-issueness differ not just with respect to the required sentence type but also in whether it is a question- or anaphora-based diagnostic. In Exps.~2, at-issueness was measured a version of the `affirmation diagnostic' in Exp.~2q and by a version of the `assent diagnostic' in Exps.~2n, 2m, and 2c. Specifically, it was measured by versions of these diagnostics in which the affirmation ({\em yes}) or assent ({\em yes, that's true}) is followed by the content to be diagnosed. In Exps.~3, at-issueness was measured by a different version of these diagnostics, namely by the version in which the affirmation or assent is followed by a denial of another salient content.

%
%In Exps.~1, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q, illustrated in (\ref{ai-exps1}a), and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c, illustrated in (\ref{ai-exps1}b). The `asking whether' diagnostic requires that the content to be diagnosed is realized by an expression in a polar question; this diagnostic was also used in \citealt{tbd-variability} Exps.~1. A speaker (Christopher in (\ref{ai-exps1}a) asks a polar question and the participant is asked whether the speaker is asking about the content to be diagnosed (here, that Julian dances salsa). If a participant takes the speaker to be asking of the content, the content is assumed to be at-issue; if a participant does not take the speaker to be asking about the content, the content is assumed to be not-at-issue. The `sure that' diagnostic requires that the content to be diagnosed is realized by an expression in declarative sentence; this diagnostic was also used in \citealt{tbd-variability} Exps.~2. A speaker utters the declarative sentence and the addressee (here, Sandy) asks {\em Are you sure?}. The speaker response with {\em Yes, I am sure that} followed by the expression that denotes the content to be diagnosed. Participants are asked whether the speaker answered the addressee's question. If a participant takes the speaker to have answered the question, the content is assumed to be at-issue; if a participant does not take the speaker to have answered the question, the content is assumed to be not-at-issue. {\bf move all this into section 1}
%
%\begin{exe}
%\ex\label{ai-exps1} At-issueness measures in Exps.~1
%\begin{xlist}
%\ex `asking whether' \\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~1q]
%\\ Question to participants: Is Christopher asking whether Julian dances salsa?
%
%\ex `sure that' \\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~1n]
%\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~1m]
%\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~1c]
%\\ Sandy: Are you sure?
%\\ Christopher: Yes, I'm sure that Julian dances salsa.
%\\ Question to participants: Did Christopher answer Sandy's question?
%
%\end{xlist}
%\end{exe}
%
%In Exps.~2, at-issueness was measured with an assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the content to be diagnosed. Participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa). The second speaker responds in the affirmative (for polar interrogatives, as shown in (\ref{ai-exps2}a)) or assents (for declaratives, as shown in (\ref{ai-exps2}b). The affirmation/assent is followed by the content to be diagnosed, here that Julian dances salsa. 
%
%\begin{exe}
%\ex\label{ai-exps2} At-issueness measures in Exps.~2
%\begin{xlist}
%\ex {\em yes}, $c$
%\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~2q]
%\\ Sandy: Yes, Julian dances salsa.
%\\ Question to participants: Does Sandy's response to Christoper sound good?
%
%\ex {\em yes, that's true}, $c$
%\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~2n]
%\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~2m]
%\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~2c]
%\\ Sandy: Yes, that's true, Julian dances salsa.
%\\ Question to participants: Does Sandy's response to Christopher sound good?
%
%\end{xlist}
%\end{exe}
%Participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance. If a participant judges the response to be acceptable, the relevant content is taken to be at-issue. If a participant judges the response to be unacceptable, the relevant content is taken to be not at-issue.
%
%In Exps.~3, at-issueness was measured with another assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the denial of another, salient content. As in the other assent diagnostic, participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa). The second speaker responds in the affirmative (for polar interrogatives, as shown in (\ref{ai-exps3}a)) or assents (for declaratives, as shown in (\ref{ai-exps3}b). The affirmation/assent is followed by the denial of another salient content, here, that Cole (didn't) discover it.
%
%\begin{exe}
%\ex\label{ai-exps3} At-issueness measures in Exps.~3
%\begin{xlist}
%\ex {\em yes}, $c$
%\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~3q]
%\\ Sandy: Yes, but Cole didn't discover it. {\bf or that?}
%\\ Question to participants: Does Sandy's response to Christoper sound good?
%
%\ex {\em yes, that's true}, $c$
%\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~3n]
%\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~3m]
%\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~3c]
%\\ Sandy: Yes, that's true, but Cole discovered / didn't discover it. {\bf or that?}
%\\ Question to participants: Does Sandy's response to Christopher sound good?
%
%\end{xlist}
%\end{exe}
%As with the other assent diagnostic, participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance. If a participant judges the response to be acceptable, the relevant content is taken to be at-issue. If a participant judges the response to be unacceptable, the relevant content is taken to be not at-issue.

@@

We recruited 250-300 participants for each of Exps.~1-3. Participants for Exp.~1q were recruited on Amazon's Mechanical Turk platform; the participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs. Participants for the remaining experiments were recruited on Prolific; these participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%. Information on the participants and the exclusion criteria is provided in Supplement \ref{a-participants}.

@@

The projection and at-issueness of the contents of 20 clauses (such as {\em Julian dances salsa}) were measured in separate blocks in Exps.~1-3. For the full set of clauses, see Supplement \ref{a-clauses}.

In the projection blocks of Exps.~1-3, target stimuli consisted of a polar question sentence, a negated sentence, a sentence with {\em perhaps}, or a conditional sentence that was uttered by a named speaker, as shown in Fig.~\ref{fig:projection-trials}. Each sentence was formed by realizing the 20 clauses as the complements of the 20 clause-embedding predicates in Fig.~\ref{fig:predicates}.\footnote{The proper names that realized the speakers, the subjects of the clause-embedding predicates, the subjects of the complement clauses, and the subjects of the consequents  were all unique. The indirect object of {\em inform} was realized by the proper name {\em Sam}. The predicates differed in the tense in which they were realized. In polar interrogatives as well as in negated and modalized declaratives, eventive predicates, like {\em discover} and {\em hear}, were realized in the past tense and stative predicates, like {\em know} and {\em be annoyed}, were realized in the present tense.  In the antecedents of conditionals, all predicates were realized in the present tense. For the consequents of the conditional sentences see Supplement \ref{a-target}.} Projection was measured using the `certain that' diagnostic \cite{djaerv-bacovcin-salt27,tbd-variability, lorson2018, mahler2020}: participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact. They gave their responses on a slider marked `no' at one end (coded as 0) and `yes' at the other (coded as 1). Greater speaker commitment to the CC should result in higher slider ratings.

@@

In the at-issueness blocks of Exps.~1-3, the 20 clauses were also realized as the complements of the 20 clause-embedding predicates in Fig.~\ref{fig:predicates} and the target stimuli realized either a polar question sentence (in Exps.~1q, 2q, and 3q), a negated sentence (in Exps.~1n, 2n, and 3n), a sentence with {\em perhaps} (in Exps.~1m, 2m, 3m), or a conditional sentence (in Exps.~1c, 2c, 3c). Sample trials of the at-issueness measures implemented in the 12 experiments are provided in Fig.~\ref{fig:ai-exps1} for Exps.~1, in Fig.~\ref{fig:ai-exps2} for Exps.~2, and in Fig.~\ref{fig:ai-exps3} for Exps.~3. Across all these measures, participants gave their responses on a slider marked `no' at one end (coded as 1) and `yes' at the other (coded as 0). Greater not-at-issueness of the CC should result in higher slider ratings.

@@

\caption{Target trials in at-issueness blocks of Exps.~2 for the complement {\em Julian dances salsa}.}\label{fig:ai-exps2}
\end{figure}

@@

To assess whether participants were attending to the task, each experiment also included six control stimuli, which were also utterances made by a named speaker. Depending on the experiment, the content of these stimuli was expected to receive either high or low certainty ratings, and either high or low ratings on the at-issueness measure; see Supplement \ref{a-control} for details and the control stimuli.

Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody say something to somebody else. Each participant saw a random set of 26 stimuli: each set contained one target utterance for each of the 20 contents (each paired with a unique clause-embedding predicates) and the same 6 (experiment-specific) control stimuli. Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block. Block order and within-block trial order was randomized. After completing the experiment, participants filled out a short optional demographic survey. To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.

@@

\section{20 clauses}\label{a-clauses}

The content of the following 20 clauses, which realized the complements of the 20 clause-embedding predicates, were investigated in Exps.~1-3:

@@

\section{Consequents for conditional target stimuli in Exps.~1c, 2c, and 3c}\label{a-consequents}

@@

The control stimuli in Exps.~1-3 were the contents of main clauses. In Exps.~1q, 2q and 3q, the control stimuli consisted of the polar questions in (1). The non-restrictive relative clauses (NRRCs), given in parentheses in (1), were included in Exps.~2q and 3q, where at-issueness was measured with an assent diagnostic. The control stimuli here consisted of two clauses (like the target stimuli), to allow the relevant speaker to assent with one of two clauses.

@@

We expected participants to give low responses on the `certain that' diagnostic for the control stimuli in (1), indicating that the speaker is not certain of the main clause content, because main clause content is hypothesized to not project out of the question. These expectations were borne out, as shown in the third column of Table~\ref{t-controls} for Exps.~1q, 2q, and 3q. We also expected participants to give low responses on the at-issueness diagnostics for the control stimuli in (1), indicating that the main clause content is at-issue. These expectations were borne out for Exps.~1q and 2q, as shown in the fourth column of Table~\ref{t-controls}, but the ratings were higher than expected for Exp.~3q. {\bf NEED TO DISCUSS THIS}

In the remaining experiments, the control stimuli consisted of the positive declarative variants of (1) given in (2). The NRRCs in parentheses were realized in Exps.~2 and 3 for the reason explained above, as well as in Exp.~1c, to make the control stimuli more similar to the target stimuli (which also consisted of two clauses, namely the antecedent and the consequent).

@@

We expected participants to give high responses on the `certain that' diagnostic for the control stimuli in (2), indicating that the speaker is certain of the main clause content, because speakers are hypothesized to be committed to the asserted main clause content. These expectations were borne out, as shown in the third column of Table~\ref{t-controls} for the `n', `m' and `c' variants of Exps.~1, 2, and 3. We expected participants to give low responses on the at-issueness diagnostics for the control stimuli in (2), indicating that the main clause content is at-issue. These expectations were borne out for the `n', `m', and `c' variants of Exps.~1, as shown in the fourth column of Table~\ref{t-controls}, but not for the `n', `m', and `c' variants of Exps.~2 and 3. {\bf NEED TO DISCUSS THIS}

\begin{table}[h!]
\centering
\begin{tabular}{r r r r l }
& &  \multicolumn{2}{c}{Mean ratings} &  \\ 
Exp. & Control stimuli & Certainty & Not-at-issueness & At-issueness measure \\ 
\hline
1q & (1) & .14 & .05  & asking whether $c$ \\
1n & (2) &  .95 & .04 & sure that $c$\\
1m & (2) & .96 & .03 & sure that $c$\\
1c & (2) with NRRC & .94  & .08 & sure that $c$\\
\hline
2q & (1) with NRRC & .18 & .07 & {\em yes}, $c$\\
2n & (2) with NRRC& .96 & .22 & {\em yes, that's true}, $c$\\
2m & (2) with NRRC& .96 & .25 & {\em yes, that's true}, $c$\\
2c & (2) with NRRC& .96 & .22 & {\em yes, that's true}, $c$\\
\hline
3q & (1) with NRRC & .17 &  .28 & {\em yes}, but $\neg c'$\\
3n & (2) with NRRC & .94 & .44 & {\em yes, that's true}, but $\neg c'$\\
3m & (2) with NRRC & .93 & .50 & {\em yes, that's true}, but $\neg c'$\\
3c & (2) with NRRC & .93 & .53 & {\em yes, that's true}, but $\neg c'$\\
\hline
\end{tabular}
\caption{Mean certainty and at-issueness ratings for control stimuli, for self-declared American English participants}\label{t-controls}
\end{table}

@@

\item `controls': Participants' data were excluded if their mean rating on the 6 main clause control items in the projection block was more than 2 sd above the group mean (in Exps.~1q, 2q, and 3q) or more than 2 sd below the group mean (in the remaining experiments). Participants' data were also excluded if their mean rating on the 6 main clause control items in the at-issueness block was more than 2 sd above the group mean (across all experiments). {\bf REVIEW THIS, MAKE SURE WE REALLY DO THIS}

@@
